---
title: "ap3650_nyc_crime_data_visualization"
author: "Anita"
date: "March 13, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      cache = TRUE)
```


```{r }
library(data.table)
library(vcdExtra)
library(extracat)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)

#fread("NYPD_Complaint_Data_Historic.csv",na.strings="",colClasses = c(PARKS_NM="c",HADEVELOPT="c"))->crime_df
#fread("NYPD_Complaint_Data_Historic.csv",na.strings="")->crime_df
crime_df <- read.csv("NYPD_Complaint_Data_Historic.csv", header=TRUE)
```

## Data Manipulations


```{r }

#Convert dates and times to correct format
crime_df$CMPLNT_FR_DT <- as.Date(crime_df$CMPLNT_FR_DT, format='%m/%d/%Y')
crime_df$CMPLNT_TO_DT <- as.Date(crime_df$CMPLNT_TO_DT, format='%m/%d/%Y')
crime_df$RPT_DT <- as.Date(crime_df$RPT_DT, format='%m/%d/%Y')

crime_df$CMPLNT_FR_TM <- as.POSIXct(crime_df$CMPLNT_FR_TM, format='%H:%M:%S')
crime_df$CMPLNT_TO_TM <- as.POSIXct(crime_df$CMPLNT_TO_TM, format='%H:%M:%S')

```
## Plots  
## Warm-up Plot :-) Bar Chart 
```{r}
  ggplot(crime_df,aes(LAW_CAT_CD)) +
    geom_bar() +
    ggtitle("Distribution of Crime Category")
```  

## Type of Offense  

```{r fig.height=10, fig.width=8}
  ggplot(crime_df,aes(PREM_TYP_DESC)) +
    geom_bar() +
    facet_wrap(~LAW_CAT_CD) +
    coord_flip() + 
    ggtitle("Crime Category Vs Place of Crime")
  

```

## Month and Time and Type of Crime

```{r fig.height = 10, fig.width=10}
  crime_df <- crime_df %>% drop_na()
  ggplot(crime_df, aes(LAW_CAT_CD)) +
  geom_bar() +
  #facet_wrap(~month(CMPLNT_FR_DT))
  #facet_wrap(~hour(CMPLNT_FR_TM))
  facet_grid(month(CMPLNT_FR_DT)~hour(CMPLNT_FR_TM))

```

## Time Series - Trend of Crime Rate 
```{r}
    crime_time <- crime_df %>%
                    filter(year(CMPLNT_FR_DT)>2005) %>%
                    group_by(Date=floor_date(CMPLNT_FR_DT, "year"),LAW_CAT_CD) %>%
                    summarize(count=n())

    ggplot(crime_time, aes(Date,count, color=LAW_CAT_CD))+
          geom_line() +
          ggtitle("Trend/Rate of Crimes in Each Category Across year")

    
    crime_time <- crime_df %>%
                    filter(year(CMPLNT_FR_DT)>2005) %>%
                    group_by(Date=floor_date(CMPLNT_FR_DT, "month"),LAW_CAT_CD) %>%
                    summarize(count=n())

    ggplot(crime_time, aes(Date,count, color=LAW_CAT_CD))+
          geom_line() +
          ggtitle("Trend/Rate of Crimes in Each Category Across year - sampled month-wise")
    
    crime_boro <- crime_df %>%
                      filter(year(CMPLNT_FR_DT)> 2005 & BORO_NM != "") %>%
                      group_by(Date=floor_date(CMPLNT_FR_DT,"month"),BORO_NM) %>%
                      summarize(count=n())
    
    ggplot(crime_boro,aes(Date,count,color=BORO_NM)) +
      geom_line() +
      ggtitle("Crime Trend over Years comparing Boroughs")

```
* Shows monthly pattern similar to Jingbo's
* Year pattern fluctuates 
* Some NM_BORO are empty
* Gaps between bororughs reduces towards later years


## length of Crime Vs Type of Crime 
```{r fig.height =20, fig.width=15}
  crime_time <- crime_df %>%  drop_na() %>%
                    filter(year(CMPLNT_FR_DT)>2005) %>%
                    mutate(delta_time = as.numeric(CMPLNT_TO_DT - CMPLNT_FR_DT)) %>%
                    group_by(OFNS_DESC, delta_time) %>%
                    summarize(count=n())
                
  ggplot(crime_time,aes(delta_time)) +
    geom_histogram(na.rm=TRUE) +
    facet_wrap(~OFNS_DESC, ncol = 4)
  
```
1. There are some cases where there might be typo on "To Date" especially year might be typo   
2. Observed larceny( grand and petite have lot of cases )  
3. There are blank OFfense category

## Atempted Crime vs Type of Crime
```{r fig.height=12, fig.width=8}
  crime_stat <- crime_df %>%
                  filter(CRM_ATPT_CPTD_CD == "ATTEMPTED" & OFNS_DESC !="") %>%
                  group_by(OFNS_DESC) %>%
                  summarize(count=n())
  ggplot(crime_stat,aes(OFNS_DESC,count)) +
      geom_col() +
      coord_flip() +
      ggtitle("Attempted Crime Status for Different Types of crime")
```

## Attempted Crime Trend 
```{r}
  crime_stat <- crime_df %>% drop_na() %>%
                  filter(CRM_ATPT_CPTD_CD=="ATTEMPTED" & year(CMPLNT_FR_DT)>2005 & LAW_CAT_CD != "") %>%
                  group_by(Date=floor_date(CMPLNT_FR_DT,"year"),LAW_CAT_CD) %>%
                  summarize(count=n())
  ggplot(crime_stat, aes(Date,count,color=LAW_CAT_CD)) +
    geom_line()
    

                  
```

## To find Top 10 Crime Categories, mosaic plots building blocks
```{r fig.height=12, fig_width = 10}
  crime_top <- crime_df %>% drop_na() %>%
                  filter(OFNS_DESC!="") %>%
                  group_by(OFNS_DESC) %>%
                  summarize(count=n())

  ggplot(crime_top, aes(reorder(OFNS_DESC,count), count)) +
    geom_col() + 
    coord_flip() +
    ggtitle(" Top Crime Categories")
  

```

## Boro, Juris, Crime Categories
```{r}
                  #group_by(BORO_NM,JURIS_DESC,OFNS_DESC) %>% 
                  #mutate(count=n()) %>%
  top_ofns   <- c("PETIT LARCENY", "HARRASSMENT 2", "CRIMINAL MISCHIEF & RELATED OF", "ASSAUALT 3 & RELATED OFFENSES", "GRAND LARCENY", "DANGEROUS DRUGS")
  crime_sort <- crime_df %>% drop_na() %>%
                  filter(BORO_NM != "", JURIS_DESC !="", (OFNS_DESC == top_ofns))%>%
                  group_by(BORO_NM,LAW_CAT_CD,OFNS_DESC) %>% 
                  summarize(Freq=n())
  
  crime_sort$TOP_OFFENSE = crime_sort$OFNS_DESC[,drop=TRUE]
  
  #doubledecker(OFNS_DESC~BORO_NM, data=crime_sort, gp = gpar(fill = c("grey90", "red")))
  mosaic(TOP_OFFENSE~BORO_NM, direction=c("v"), labeling=labeling_border(rot_labels=c(45,0,0, 0)), crime_sort)
  
  
  #doubledecker(TOP_OFFENSE~BORO_NM, data=crime_sort)
  ggplot(crime_sort, aes(TOP_OFFENSE,Freq, fill=BORO_NM)) +
    geom_col() +
    facet_grid(LAW_CAT_CD~ BORO_NM) +
    coord_flip()
```

** The above plot shows something surprising, the categories are not standard, need to research more. For example, dangerous drugs is under Felony as well as Misdemeanor!! **


```{r}
  ggplot(crime_sort, aes(TOP_OFFENSE,Freq)) +
           geom_col() +
           facet_wrap(~BORO_NM) + 
           coord_flip()
                         
```

* I tired indivial Crime Types, the colors were too confusing as lot of categories


```{r fig.height=16}
  crime_parks <- tbl_df(crime_df) %>% drop_na() %>%
                  filter(BORO_NM!="",PARKS_NM!="",LAW_CAT_CD!="") %>%
                  group_by(BORO_NM,PARKS_NM,LAW_CAT_CD) %>%
                  summarize(count=n())

  #crime_parks <- crime_parks %>%
  #                arrange(desc(count))
  
 crime_pk <- crime_parks %>% 
                 group_by(BORO_NM) %>%
                 top_n(n=10, wt=count)
 
  crime_parks_1 <- crime_df %>% drop_na() %>%
                  filter(BORO_NM!="",PARKS_NM!="",OFNS_DESC!="") %>%
                  group_by(BORO_NM,PARKS_NM,OFNS_DESC) %>%
                  summarize(count=n())
 
 crime_pk_1 <- crime_parks_1 %>% 
                 group_by(BORO_NM) %>%
                 top_n(n=10, wt=count)
 
 
  ggplot(crime_pk_1 ,aes(PARKS_NM, count, fill=OFNS_DESC)) +
    geom_col() + 
    facet_wrap(~BORO_NM, ncol=1, scales="free_y") +
    coord_flip()
```

## trial on ggmap
```{r}
  library(ggmap)
  
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
