---
title: "=====Part 4===== Main Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE, 
                      cache = TRUE)
```

```{r cars}
#if you have not please install data.table package before run the codes below
#install.packages(data.table)
library(zoo)
library(vcd)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(vcdExtra)
library(gridExtra)
library(tidyverse)
library(data.table)
fread("NYPD_Complaint_Data_Historic.csv",na.strings="",colClasses = c(PARKS_NM="c",HADEVELOPT="c"))->df
```
\
\
```{r}
#picking non-missing CMPLNT_FR_DT and convert to Date and filter only those after "2006-01-01", 5560408 obs.
df%>%select(CMPLNT_FR_DT,LAW_CAT_CD)%>%filter(!is.na(CMPLNT_FR_DT))%>%mutate(CMPLNT_FR_DT=as.Date(CMPLNT_FR_DT,format='%m/%d/%Y'))%>%filter(CMPLNT_FR_DT>=as.Date("2006-01-01"))->df_Date
```

```{r}
#time series of daily frequency of 3 crime categories 2006-2016
df_Date%>%group_by(CMPLNT_FR_DT,LAW_CAT_CD)%>%dplyr::summarise(count=n())%>%ungroup()%>%group_by(LAW_CAT_CD)%>%mutate(mon_mean=rollmean(count,30,fill=NA))%>%ungroup()->byDateLawMean

#daily rate
byDateLawMean%>%ggplot()+
  geom_line(aes(CMPLNT_FR_DT,count,color=LAW_CAT_CD))+
  geom_line(aes(CMPLNT_FR_DT,mon_mean,group=LAW_CAT_CD))+
  ggtitle("Daily Crime Frequency since 2006 with 30-day running mean")+
  labs(x="Date",y="Frequency")+theme(legend.title=element_blank())+geom_line(aes(CMPLNT_FR_DT,count*0+1150))

#Top 9 daily rate falls on Jan 1.
byDateLawMean%>%arrange(desc(count))%>%filter(count>=count[9])%>%
  mutate(CMPLNT_FR_DT=as.factor(CMPLNT_FR_DT))%>%
  ggplot(aes(forcats::fct_reorder(CMPLNT_FR_DT, count),count))+geom_bar(stat="identity")+coord_flip()+ylab("Top 9 Daily Crime Frequency")+xlab("Date")
```

* The crime frequency is decreasing over the years this is because lots of cases occurred over the years haven't reported yet. 
* There are obvious annual variation/cycle. 30-day running mean shows the cycle clearly.
* There are spikes in the misdemeanor category. The top 9 dates with high frequency are shown in the barchart. They are on January 1 on almost each year from 2006-2016 except 2015 which is actually very close behind. These cases seemed like mistakening assigned an occurrence date as January 1 since by examining the relationships between RPT_DT,CMPLNT_FR_DT and CMPLNT_TO_DT, they don't seem make much sense comparing with others.
\
\
```{r}
#frequency by month
df_Date%>%mutate(Month=as.character(month(CMPLNT_FR_DT)))%>%group_by(Month,LAW_CAT_CD)%>%dplyr::summarise(CntByMon=n())->byDateLaw_mon

byDateLaw_mon%>%mutate(Days=rep(31,3))%>%mutate(Days=ifelse(Month=="2",28,Days))%>%mutate(Days=ifelse(Month %in% c("4","6","9","11"),30,Days))->byDateLaw_mon
byDateLaw_mon%>%ggplot(aes(fct_relevel(Month,"10","11","12",after=9),CntByMon/Days))+geom_bar(stat="identity")+coord_flip()+ylab("Crime Frequency (Monthly Mean)")+facet_wrap(~LAW_CAT_CD,scales="free_x")+xlab("Month")->p1
```

```{r}
#frequency by day
df_Date%>%mutate(Day=as.factor(format(CMPLNT_FR_DT,"%d")))%>%group_by(Day,LAW_CAT_CD)%>%dplyr::summarise(CntByDay=n())->byDateLaw_day

#Day1-28 has the same total cnts=11yrs*12cnts/yr
#Day 29 cnts=11yrs*11cnts/yr+3cnts (leap yrs)
#Day 30 cnts=11*11; Day 31 cnts=7*11
byDateLaw_day%>%mutate(cnts=rep(12*11,3))%>%mutate(cnts=ifelse(Day=="29",11*11+3,cnts))%>%mutate(cnts=ifelse(Day=="30",11*11,cnts))%>%mutate(cnts=ifelse(Day=="31",7*11,cnts))->byDateLaw_day

byDateLaw_day%>%ggplot(aes(Day,CntByDay/cnts))+geom_bar(stat="identity")+coord_flip()+ylab("Crime Frequency (Daily Mean)")+facet_wrap(~LAW_CAT_CD,scales="free_x")+xlab("Day of Month")->p2
```


```{r}
#frequency by weekday
df_Date%>%mutate(Wkday=as.factor(weekdays(CMPLNT_FR_DT,abbreviate=TRUE)))%>%group_by(Wkday,LAW_CAT_CD)%>%dplyr::summarise(CntByWkday=n())->byDateLaw_wkday

byDateLaw_wkday%>%ggplot(aes(fct_relevel(Wkday,"Mon","Tue","Wed","Thu","Fri","Sat","Sun"),CntByWkday))+geom_bar(stat="identity")+coord_flip()+ylab("Crime Frequency")+facet_wrap(~LAW_CAT_CD,scales="free_x")+xlab("Day of Week")->p3
```

```{r}
#picking non-missing CMPLNT_FR_TM
df%>%filter(!is.na(CMPLNT_FR_TM))%>%mutate(CMPLNT_FR_DT=as.Date(CMPLNT_FR_DT,format='%m/%d/%Y'))%>%filter(CMPLNT_FR_DT>=as.Date("2006-01-01"))->df_FRTM

#Frequency by hour of day, combining hour 00 and hour 24 into hour 00
df_FRTM%>%mutate(Hour=as.factor(substr(CMPLNT_FR_TM,1,2)))%>%group_by(Hour,LAW_CAT_CD)%>%dplyr::summarise(CntByHour=n())->byDateLaw_hour
byDateLaw_hour$Hour[byDateLaw_hour$Hour=="24"]<-"00"
byDateLaw_hour$Hour<-factor(byDateLaw_hour$Hour)

byDateLaw_hour%>%ggplot(aes(Hour,CntByHour))+geom_bar(stat="identity")+coord_flip()+ylab("Crime Frequency")+facet_wrap(~LAW_CAT_CD,scales="free_x")+xlab("Hour of Day")->p4
```

```{r fig.height=4,fig,width=6}
grid.arrange(p1,p2,p3,p4,nrow=2)
```
\
\
* Indeed by barcharting over the months,we see Jun.-Oct. is a high crime season.
* The fake January increasing was due to the errors in the records. 
\
\
* The spike in Janurary is consistent with the analysis above. 
* There seemed having a tendency of rounding every 5 day.
\
\
* Violation is low during weekends but same during weekdays.
* Felony and misdemeanor is high on Friday but low on Sunday nad Monday.
\
\
* There is obvious day cycle in the crime occurrence. Early morning has the least crime occurrence while later afternoon has the most crime occurrence.
\
\
```{r fig.height=5,fig.width=5}
#how the different crime types (OFNS_DESC) associated with different places (a heatmap)
#first filling the missing OFNS_DESC infered from KY_CD
df%>%select(KY_CD,OFNS_DESC)%>%group_by(KY_CD)%>%
  dplyr::summarise(desc=paste(unique(OFNS_DESC),collapse=","))%>%
  mutate(KY_CD=as.factor(KY_CD))%>%arrange(desc)->match_code_desc

df%>%
  select(KY_CD,PREM_TYP_DESC)%>%
  filter(!is.na(PREM_TYP_DESC))%>%
  mutate(KY_CD=as.factor(KY_CD))%>%
  group_by(KY_CD,PREM_TYP_DESC)%>%dplyr::summarise(count=n())%>%mutate(pct=count/sum(count))->byKYbyPREM

#merging to get OFNS_DESC vs PREM_TYP_DESC correspondence
merge(byKYbyPREM, match_code_desc, by.x='KY_CD', by.y='KY_CD')->byKYbyPREM_match

byKYbyPREM_match%>%group_by(desc)%>%dplyr::summarise(mean=mean(count),na.rm=TRUE)%>%arrange(desc(mean))->desc_desc_cnt
byKYbyPREM_match%>%group_by(PREM_TYP_DESC)%>%dplyr::summarise(mean=mean(count),na.rm=TRUE)%>%arrange(desc(mean))->PREM_desc_cnt

byKYbyPREM_match%>%
  ggplot(aes(fct_relevel(as.factor(desc),as.character(desc_desc_cnt$desc[sort(desc_desc_cnt$mean,index.return=TRUE,decreasing=TRUE)$ix])),
  fct_relevel(as.factor(PREM_TYP_DESC),as.character(PREM_desc_cnt$PREM_TYP_DESC[sort(PREM_desc_cnt$mean,index.return=TRUE,decreasing=TRUE)$ix])),fill=pct))+scale_fill_gradientn(colors=c("red","orange","yellow","green","blue","violet"),na.value="black")+
  scale_x_discrete(label=function(x) abbreviate(x, minlength=20))+coord_flip()+
    geom_tile(color="white",size=0.25)+theme(axis.text.x = element_text(size=5,angle = 90, hjust = 1),axis.text.y=element_text(size=5))+ylab("Premises")+xlab("OFNS_DESC")
```
\
\
* Doesn't seem having association between crime types and premises.
\
\
```{r fig.height=5,fig.width=5}
#how the different crime types associated with time using heatmap
df%>%
  select(KY_CD,CMPLNT_FR_TM)%>%
  filter(!is.na(CMPLNT_FR_TM))%>%
  mutate(KY_CD=as.factor(KY_CD))%>%
  mutate(Hour=as.factor(substr(CMPLNT_FR_TM,1,2)))%>%
  group_by(KY_CD,Hour)%>%dplyr::summarise(count=n())%>%mutate(pct=count/sum(count))->byKYbyFRTM
#combining hour 00 and hour 24 into hour 00
byKYbyFRTM$Hour[byKYbyFRTM$Hour=="24"]<-"00"
byKYbyFRTM$Hour<-factor(byKYbyFRTM$Hour)

#merging to get OFNS_DESC vs CMPLNT_FR_TM correspondence
merge(byKYbyFRTM, match_code_desc, by.x='KY_CD', by.y='KY_CD')->byKYbyFRTM_match

byKYbyFRTM_match%>%group_by(desc)%>%dplyr::summarise(mean=mean(count),na.rm=TRUE)%>%arrange(desc(mean))->desc2_desc_cnt
byKYbyFRTM_match%>%group_by(Hour)%>%dplyr::summarise(mean=mean(count),na.rm=TRUE)%>%arrange(desc(mean))->Hour_desc_cnt

byKYbyFRTM_match%>%ggplot(aes(
  fct_relevel(as.factor(desc),as.character(desc2_desc_cnt$desc[sort(desc2_desc_cnt$mean,index.return=TRUE,decreasing=TRUE)$ix])),
  Hour,fill=pct))+scale_fill_gradientn(colors=c("red","orange","yellow","green","blue","violet"),na.value="black")+
  scale_x_discrete(label=function(x) abbreviate(x, minlength=20))+coord_flip()+
    geom_tile(color="white",size=0.25)+theme(axis.text.x = element_text(size=5,angle = 90, hjust = 1),axis.text.y=element_text(size=5))+ylab("Hour")+xlab("OFNS_DESC")
```
\
\
* Do we see any association between time and certain crime? Do see some high density around middle up right area, which is consistent with the barcharting daily cycle.
\
\
```{r fig.height=4}
#matching Pct with Boro
df%>%select(ADDR_PCT_CD,BORO_NM)%>%filter(!is.na(ADDR_PCT_CD) & !is.na(BORO_NM))%>%group_by(ADDR_PCT_CD)%>%
  dplyr::summarise(boro=paste(unique(BORO_NM),collapse=","))%>%
  mutate(ADDR_PCT_CD=as.factor(ADDR_PCT_CD))%>%arrange(boro)->match_pct_boro

df %>% select(LAW_CAT_CD,ADDR_PCT_CD)%>%group_by(LAW_CAT_CD,ADDR_PCT_CD)%>%
  drop_na()%>%dplyr::summarize(count = n())%>%ungroup()->df_pct
merge(df_pct,match_pct_boro,by.x="ADDR_PCT_CD",by.y="ADDR_PCT_CD")%>%arrange(desc(count))->df_pbl

df_pbl%>%ggplot(aes(reorder(ADDR_PCT_CD, count), count,fill=boro)) + geom_bar(stat = "identity") + xlab("Precint Number") + ggtitle("Incidents by Precinct") + coord_flip()+scale_fill_manual(values = c("dark green", "orange","blue","green","red","brown","violet","yellow","navy","grey","pink","light blue"))
```

* Just to see if top precints occured in certain borough. Same as Rich's so if we are going to select this, just add color to Rich's. So we gain more information in one figure. Note, precint is not one-to-one correspondence to borough! Both claim to be where the incident occurred as described in doc.
\
\