---
title: "Plot Ideas for Meteorological and Lunar Relationships"
author: "Rich Lavery"
date: "April 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE, 
                      cache = TRUE)
```


### Read In Dataset from .csv file and merge datasets on weather and lunar cycles

```{r, message = FALSE, fig.width = 10, fig.height = 8}
#Note the "NYPD_Complaint_Data_Historic.csv" should be in the working directory when running this file.
#Note the "nyc_weather_data.csv" file should be in the working directory when running this file.
#Note the "nyc_moon_data.csv" file should be in the working directory when running this file.

library(dplyr)
library(tidyr)
library(ggplot2)
library(vcdExtra)
library(data.table)

########################################################################################
var_names <- c("Id", "DateStart", "TimeStart", "DateEnd", "TimeEnd", "DateReport", "ClassCode", "OffenseDesc", 
               "IntClassCode", "IntOffenseDesc", "AtptCptdStatus", "Level", "Jurisdiction", "Boro", "Pct", "LocOfOccr", "PremDesc", 
               "ParkName", "HousingDevName", "XCoord", "YCoord", "Lat", "Long", "Lat_Long")

crime_df <- fread("NYPD_Complaint_Data_Historic.csv",na.strings="", col.names = var_names, stringsAsFactors = TRUE)

#Convert dates and times to correct format
crime_df$DateStart <- as.Date(crime_df$DateStart, format='%m/%d/%Y')
crime_df$DateEnd <- as.Date(crime_df$DateEnd, format='%m/%d/%Y')
crime_df$DateReport <- as.Date(crime_df$DateReport, format='%m/%d/%Y')

########################################################################################
#Read in weather data from file
weather_select = c("DATE", "AWND", "PRCP", "SNOW", "TMAX")
weather_data <- fread("nyc_weather_data.csv", na.strings="", select = weather_select, stringsAsFactors = FALSE)
weather_data$DATE <- as.Date(weather_data$DATE)
weather_data$AWND <- as.numeric(weather_data$AWND)
weather_data$PRCP <- as.numeric(weather_data$PRCP)
weather_data$SNOW <- as.numeric(weather_data$SNOW)
weather_data$TMAX <- as.numeric(weather_data$TMAX)

#Merge the data together
crime_df <- crime_df[weather_data, on=.(DateStart = DATE)]

###############################################################################
#Read in moon phase data
moon_data <- fread("nyc_moon_data.csv", na.strings="", select = c("date", "phase"), stringsAsFactors = FALSE)
moon_data$date <- as.Date(moon_data$date, format='%m/%d/%Y')
moon_data$phase <- as.factor(moon_data$phase)
full_moon_data <- moon_data %>% filter(phase == "Full Moon")

#Merge the moon phase data into the main data frame
crime_df <- crime_df %>% left_join(moon_data, by = c("DateStart" = "date"))

#Generate violent crime dataframe
#filter for violent crime
violent_crime_df <- crime_df %>% filter(OffenseDesc == "ASSAULT 3 & RELATED OFFENSES" | 
                                          OffenseDesc == "FELONY ASSAULT" | 
                                          OffenseDesc == "MURDER & NON-NEGL. MANSLAUGHTER" |  
                                          OffenseDesc == "RAPE" |
                                          OffenseDesc == "ROBBERY" |
                                          IntOffenseDesc == "AGGRAVATED SEXUAL ASBUSE" |
                                          IntOffenseDesc == "ASSAULT 2,1,UNCLASSIFIED" |
                                          IntOffenseDesc == "ASSAULT 3" |
                                          IntOffenseDesc == "RAPE 1" |
                                          IntOffenseDesc == "ROBBERY,OPEN AREA UNCLASSIFIED" |
                                          IntOffenseDesc == "SEXUAL ABUSE" |
                                          IntOffenseDesc == "SEXUAL ABUSE 3,2")

```

## Investigate the effects of precipitation on crime.
### How does the average number of reported crimes on days with more than 1/2 inch of precipitation compare to the average number of crimes reported on days with no precipitation?

```{r, fig.width = 10, fig.height = 8}
#Generate scatterplot of crime vs precipitiation
rain_summary_per_day <- crime_df %>% group_by(DateStart, Level) %>% summarize(Count = n()) %>% drop_na()
#append weather data
rain_summary_per_day <- rain_summary_per_day %>% left_join(weather_data, by = c("DateStart" = "DATE")) %>% select(DateStart, Level, Count, PRCP)
#Scatter plot of daily crimes vs. precipitation level

#Filter on Level of crime and generate linear model for each
# linear model: Felonies
f_df_rain <- rain_summary_per_day %>% filter(Level=="FELONY")
flm_rain <- lm(Count~PRCP, f_df_rain)
# linear model: Misdemeanors
m_df_rain <- rain_summary_per_day %>% filter(Level=="MISDEMEANOR")
mlm_rain <- lm(Count~PRCP, m_df_rain)
# linear model: Violation
v_df_rain <- rain_summary_per_day %>% filter(Level=="VIOLATION")
vlm_rain <- lm(Count~PRCP, v_df_rain)

#Plot data for all three crime levels vs precipitation with linear model results
ggplot(rain_summary_per_day, aes(x=PRCP, y=Count, color=Level)) + 
  geom_point(alpha=0.3) +
  geom_abline(slope=flm_rain[["coefficients"]][["PRCP"]],intercept=flm_rain[["coefficients"]][["(Intercept)"]]) +
  annotate("text", x= 3, y=420, label=paste0("y=",round(flm_rain[["coefficients"]][["PRCP"]],2),"x+",round(flm_rain[["coefficients"]][["(Intercept)"]],0))) +
  geom_abline(slope=mlm_rain[["coefficients"]][["PRCP"]],intercept=mlm_rain[["coefficients"]][["(Intercept)"]]) +
  annotate("text", x= 3, y=750, label=paste0("y=",round(mlm_rain[["coefficients"]][["PRCP"]],2),"x+",round(mlm_rain[["coefficients"]][["(Intercept)"]],0))) +
  geom_abline(slope=vlm_rain[["coefficients"]][["PRCP"]],intercept=vlm_rain[["coefficients"]][["(Intercept)"]]) +
  annotate("text", x= 3, y=180, label=paste0("y=",round(vlm_rain[["coefficients"]][["PRCP"]],2),"x+",round(vlm_rain[["coefficients"]][["(Intercept)"]],0))) +
  labs(x = "Precipitation [inches]", y = "Daily Crime Incident Count", title = "Daily Crime Counts vs. Precipitation by Level of Crime with Linear Models")

#Plot data only for violent crimes vs precipitation with linear model result
#Generate scatterplot of crime vs precipitiation
vc_df_rain <- violent_crime_df %>% group_by(DateStart) %>% summarize(Count = n()) %>% drop_na()
#append weather data
vc_df_rain <- vc_df_rain %>% left_join(weather_data, by = c("DateStart" = "DATE")) %>% select(DateStart, Count, PRCP)
#Scatter plot of daily crimes vs. precipitation level

#Generate linear model for Violent Crime vs. Precipitation
vclm_rain <- lm(Count~PRCP, vc_df_rain)

ggplot(vc_df_rain, aes(x=PRCP, y=Count)) + 
  geom_point(alpha=0.3) +
  geom_abline(slope=vclm_rain[["coefficients"]][["PRCP"]],intercept=vclm_rain[["coefficients"]][["(Intercept)"]]) +
  annotate("text", x= 3.5, y=240, label=paste0("y=",round(vclm_rain[["coefficients"]][["PRCP"]],2),"x+",round(vclm_rain[["coefficients"]][["(Intercept)"]],0))) +
  labs(x = "Precipitation [inches]", y = "Daily Violent Crime Incident Count", title = "Daily Violet Crime Counts vs. Precipitation with Linear Models")
```

### We can clearly see a reduction in the average number of crimes reported on days with significant precipitation.  This is not very surprising as we would expect less people to be out on days with bad weather.  The effect of increasing precipitation on misdemeanors appears to be the most significant, with a lesser impact on felonies and violations.  When looking at the same analysis after filtering only on Violent Crimes, we see a relationship very similar to that of felonies, which is what we would expect considering most violent crimes are felonies.

## Investigate the effects of a full moon on criminal activity.
### There is an "old wives tale" about how a full moon tends to make people more erratic and violent.  This partially explains the etymology of the word "lunatic."  We are looking for evidence of this theory by merging historical data on the occurance of a full moon with our crime dataset.

```{r, fig.width = 10, fig.height = 8}
moon_summary <- crime_df %>% 
  filter(phase == "Full Moon" | phase == "New Moon" | phase == "First Quarter" | phase == "Last Quarter") %>% 
  group_by(DateStart, phase) %>% summarize(Count = n()) %>% drop_na()
moon_avg_crime <- moon_summary %>% group_by(phase) %>% summarize(Avg_Count = weighted.mean(Count))
moon_total_crime <- moon_summary %>% group_by(phase) %>% summarize(Total_Count = sum(Count))
moon_phase_total_count <- sum(moon_total_crime$Total_Count)
moon_total_crime <- moon_total_crime %>% mutate(Pct = Total_Count/moon_phase_total_count)

#Create a pie chart
blank_theme <- theme_minimal()+
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=14, face="bold")
  )

ggplot(data = moon_total_crime, aes(x="", y = Total_Count, fill = phase)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y", start=0) +
  blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(label = scales::percent(Pct)), position = position_stack(vjust = 0.5)) + 
  ggtitle("Moon Phase vs. Crime Count Analysis")

#Generate the same analysis based on Violent Crimes
vc_moon_summary <- violent_crime_df %>% 
  filter(phase == "Full Moon" | phase == "New Moon" | phase == "First Quarter" | phase == "Last Quarter") %>% 
  group_by(DateStart, phase) %>% summarize(Count = n()) %>% drop_na()
vc_moon_avg_crime <- vc_moon_summary %>% group_by(phase) %>% summarize(Avg_Count = weighted.mean(Count))
vc_moon_total_crime <- vc_moon_summary %>% group_by(phase) %>% summarize(Total_Count = sum(Count))
vc_moon_phase_total_count <- sum(vc_moon_total_crime$Total_Count)
vc_moon_total_crime <- vc_moon_total_crime %>% mutate(Pct = Total_Count/vc_moon_phase_total_count)

ggplot(data = vc_moon_total_crime, aes(x="", y = Total_Count, fill = phase)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y", start=0) +
  blank_theme +
  theme(axis.text.x=element_blank()) +
  geom_text(aes(label = scales::percent(Pct)), position = position_stack(vjust = 0.5)) + 
  ggtitle("Moon Phase vs. Violent Crime Count Analysis")
```

### We can see from the data that crime does not appear to be affected significantly on days with a full moon.  We can conclude that there is no statistical evidence to support the "old wives tale" about lunatics.

## How does the quantity of crime reports vary across the different precincts?

```{r, fig.width = 10, fig.height = 8}
#Plot histogram of incidents by police precinct
data_pct <- crime_df %>% group_by(Pct) %>% summarize(count = n()) %>% drop_na()
ggplot(data_pct, aes(reorder(Pct, count), count)) + geom_bar(stat = "identity") + xlab("Precint Number") + ggtitle("Incidents by Precinct") + coord_flip()
```


