---
title: "Rrr2149_Analysis"
author: "Rashmi"
date: "4/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This covers high level yearly symbiotic relationship between unemployment and NYC Crime Rate . 

Preparing the data for yearly crime and avg unempleoyment count.

```{r cars}
library(dplyr)
library(tidyr)
library(ggplot2)
library(vcdExtra)
library(data.table)


var_names <- c("Id", "DateStart", "TimeStart", "DateEnd", "TimeEnd", "DateReport", "ClassCode", "OffenseDesc", 
               "IntClassCode", "IntOffenseDesc", "AtptCptdStatus", "Level", "Jurisdiction", "Boro", "Pct", "LocOfOccr", "PremDesc", 
               "ParkName", "HousingDevName", "XCoord", "YCoord", "Lat", "Long", "Lat_Long")

crime_df <- fread("~/Downloads/NYPD_Complaint_Data_Historic.csv",na.strings="", col.names = var_names, stringsAsFactors = TRUE)


#Convert dates and times to correct format
crime_df$DateStart <- as.Date(crime_df$DateStart, format='%m/%d/%Y')
crime_df$DateEnd <- as.Date(crime_df$DateEnd, format='%m/%d/%Y')
crime_df$DateReport <- as.Date(crime_df$DateReport, format='%m/%d/%Y')
crime_df_master <-crime_df

########################################################################################
#Read in UnEmployment data from file

UnEmployment_data <- fread("~/Desktop/Exploratory DataAnalysis/CrimeDataNYC/Data_Files/unempCSV.csv", na.strings="", stringsAsFactors = FALSE)

UnEmployment_data$year <-as.character(UnEmployment_data$Year)
names(UnEmployment_data)[names(UnEmployment_data) == 'Ann Avg'] <- 'AnnAvg'
#UnEmployment_data
UnEmployment_data<-UnEmployment_data %>% select("year","AnnAvg")%>% drop_na()


crime_df%>%mutate(year=as.character(year(DateReport)))%>%group_by(year)%>%dplyr::summarise(CntByYear=n())%>% drop_na()->crime_df


crime_df <- merge(x=crime_df,y=UnEmployment_data,by.x="year",by.y="year")
crime_df



```

Lets prepare the yearly unemployment vs crime ratio  .
On a close look this can be observed that  ther is a  slight increase in the ratio  starting from 2006 though 2008 followed by there is a huge increase  past 2009 and the trend continue with minimal variation  untill 2013. the ratio unemployment vs crime is again decreaing pst 2013.

Relationship between Crime and Un-Employment

```{r}
### Relationship between max temp and crime volume
# set up the data by day and Level
crime_df <-crime_df %>% group_by(year) %>% summarize(CrimeCount=sum(CntByYear),AnnAvg,unEmpCrimeRatio =((AnnAvg)/sum(CntByYear)))
# plot it -- well, plot it later after the linear models are run so we can see the linear slopes
crime_df
```

Lets study the relationship more closely via below two plots.  
2006/2007/2008  : Significant low unemployment rate, and the significantly high crime rate

2009/2010/2011 : High un-employment rate Vs moderate crime

2013 onwards the relationship between growth of unemployment and crime are mutual.


```{r}
require(gridExtra)
x1 =ggplot(crime_df, aes(x=year,y=AnnAvg)) +
  geom_point( alpha=0.3)+
ylab("Annual Unemployment")+
  
x2=  ggplot(crime_df, aes(x=year,y=CrimeCount)) +
  geom_point( alpha=0.3)+xlab("crime count over 10 years")+
ylab(" Yearly Crime count ")+
 




grid.arrange(x1,x2,nrow=1)
 ggplot(crime_df,aes(x=CrimeCount,y=AnnAvg))+geom_point(alpha=0.3)+ylab(" Average yearly Unemployment ")+
 



```


The below study shows there  has been a variation in unemployemnt partten, this cuould be due to the fact that the USA East coast economy is influnced by unempleoyment which directly triggers the crime rate.
late 2008 thrugh late 2009/early 2010 the employment was gretly influence by the recession as the economy staratd picking up slowly , employment started getting better shape and that is one of the potential factors to help decrease in crime rate. Better the economy less Crime.

```{r}
ggplot(crime_df, aes(x=year, y=unEmpCrimeRatio)) + 
  geom_point(alpha=0.5) +
  ggtitle(" Yearly crime Vs ")
```


```{r}

library(tidyr)
crime_df_master

crime_df_master%>%mutate(year=as.character(year(DateReport)))%>%group_by(year,Level)%>%dplyr::summarise(CntByYear=n())%>% drop_na()->crime_df_master1

crime_df_master1 <- spread(crime_df_master1,Level,CntByYear)
crime_df_master1
write.csv(crime_df_master1, "crime_master.csv")

```

